---
title: "xxx"
author: "Jaclyn A Coate"
date: "2/9/2020"
output: html_document
---


#### Checking for Multicollinearity
### Numerical v Numerical Multicollinearity
- Multicollinearity will weaken the model
  - number_of_reviews and reviews_per_month are correlated at 55%
    - Removing reviews_per_month

```{r multicollinearity1}
corrNYC <- nyc3
#Table numeric variables
corrNYCTable <- corrNYC %>% keep(is.numeric) %>% cor %>% view
#Plot numeric variables v numeric variables
corrNYC %>% keep(is.numeric) %>% cor %>% corrplot("upper", addCoef.col = "white", number.digits = 2, number.cex = 0.5, method="square", order="hclust", tl.srt=45, tl.cex = 0.8)
invisible(view(corrNYCTable))
```

```{r attrition: removing highly correlated variable}
#Removing reviews_per_month due to high correlation of is and number_of_reviews
nyc4 <- select(nyc3, -c("reviews_per_month"))
```

### Examining VIFs for Multicollinearity of No Transformation
- The below results show us there is no need to remove any variables

```{r examining VIFs}
full.model<-lm(price~.,data=nyc4)  # . means all variable not mpg
vif(full.model)[,3]^2

alias(lm(price~.,data=nyc4))
```

#### Assumptions of Model

```{r}
par(mfrow=c(2,2))
full.model<-lm(price~.,data=nyc4)
plot(full.model)
```

#### Examining VIFs of Log Price Variable

```{r exmaining log-linear VIFs}
log.depend.model<-lm(lprice~.,data=log.nyc)  # . means all variable not mpg
vif(log.depend.model)[,3]^2

alias(lm(lprice~.,data=log.nyc))
```

#### Assumptions of Log-Linear Model

```{r}
par(mfrow=c(2,2))
log.depend.model<-lm(lprice~.,data=log.nyc)
plot(log.depend.model)
```

#### Examining VIFs of Log-Log Model

```{r examining log-log VIFs}
log.indep.model<-lm(lprice~.,data=log.indep.nyc)  # . means all variable not mpg
vif(log.indep.model)[,3]^2

alias(lm(lprice~.,data=log.indep.nyc))
```

#### Assumptions of Log-Log Model

```{r}
par(mfrow=c(2,2))
log.indep.model<-lm(lprice~.,data=log.indep.nyc)
plot(log.indep.model)
```

#### Assumptions of Binned Data

```{r}
par(mfrow=c(2,2))
plot(nyc.bin.model)
```